Metadata control document

### 2023-05-25 metadata/metadata_control.Rmd
Fully switched to metadata v2

library(tidyverse)
library(cmap4r)
cmap_api <- "6acd63a0-4c63-11ea-aa09-71d9d763e28d"
set_authorization(cmap_key = cmap_api)
library(fuzzyjoin)
library(ggmap)
register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
library(ggrepel)
library(gridExtra)  # For grid.arrange
library(data.table) # For fast indexing of seaflow data
library(RaMS)       # For MS file timestamps
library(scales)     # For "muted" function
library(readxl)     # For "readxl"

@@

if(!dir.exists("raw_data")){
  dir.create("raw_data")
}
if(!dir.exists("made_data")){
  dir.create("made_data")
}
if(!dir.exists("figures")){
  dir.create("figures")
}

options(pillar.sigfig = 7)

toDecDeg <- function(deg, min, dir){
  num <- deg+min/60
  if(dir%in%c("S", "W")){
    num*-1
  } else {
    num
  }
}
```

## CTD data (cruise, station, time, lat, lon, cast, SLA, DCM depth)

Cruise is manually entered, station/time/lat/lon/cast come from the SCOPE website cruise
summaries, SLA is pulled down from CMAP, and DCM depth was manually found by looking
at the cruise logbook.

Also now includes data from .gof files for exact depth

```{r station_metadata}
# DCM depths are entered manually ----
if(!file.exists("raw_data/all_DCMs.csv")){
  falkor_DCMs <- data.frame(
    station=c(62, 64, 77, 80),
    cast=1,
    DCM_depth=c(98, 115, 125, 120)
  )
  transect_DCMs <- data.frame(
    station = c(4:13, 13, 14),
    cast=c(1,1,3,1,1,2,2,2,1,2,3,2),
    DCM_depth=c(100, 122, 119, 133, 122, 110, 110, 110, 103, 112, 112, 105)
  )
  center_DCMs <- data.frame(
    station = paste0("L", c(1,1,2,2)),
    cast=c(32, 33, 1, 2),
    DCM_depth=c(108, 104, 123, 123)
  )
  all_DCMs <- rbind(falkor_DCMs, transect_DCMs, center_DCMs)

  write.csv(all_DCMs, "raw_data/all_DCMs.csv", row.names = FALSE)
}

# SCOPE CTD data consists of station, cast, time, max_depth, bottles, lat, and lon ----
getScopeCTD <- function(url){
  col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
  read_table(url, skip = 3, col_names = col_names, col_types = cols(
    station = col_character(),
    cast = col_double(),
    month = col_character(),
    day = col_character(),
    year = col_double(),
    time = col_time(format = ""),
    lat_deg = col_double(),
    lat_min = col_double(),
    lat_dir = col_character(),
    lon_deg = col_double(),
    lon_min = col_double(),
    lon_dir = col_character(),
    max_depth = col_double(),
    bottles = col_double())) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}

if(!file.exists("raw_data/falkor_SCOPEweb_ctd_summary.csv")){
  # Data source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/FK_CTDsummary_current.txt
  falkor_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "SCOPEcore/FK_CTDsummary_current.txt")
  falkor_SCOPEweb_ctd_summary <- getScopeCTD(falkor_SCOPEweb_ctd_summary_url)
  falkor_SCOPEweb_ctd_summary <- filter(falkor_SCOPEweb_ctd_summary, time > "2018-03-23 04:01:51")
  write.csv(falkor_SCOPEweb_ctd_summary, 
            file = "raw_data/falkor_SCOPEweb_ctd_summary.csv", 
            row.names = FALSE)
}
if(!file.exists("raw_data/meso_SCOPEweb_ctd_summary.csv")){
  # Data url source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/MS_CTDsummary_current.txt
  meso_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
                                 "@scope.soest.hawaii.edu/collaborators/datainventory/Data",
                                 "/SCOPEcore/MS_CTDsummary_current.txt")
  meso_SCOPEweb_ctd_summary <- getScopeCTD(meso_SCOPEweb_ctd_summary_url)
  write.csv(meso_SCOPEweb_ctd_summary, 
            file = "raw_data/meso_SCOPEweb_ctd_summary.csv", 
            row.names = FALSE)
}

# SLA data is pulled down from CMAP for Falkor, FTP for meso ----
# SLA units are meters
if(!file.exists("raw_data/falkor_SLA.csv")){
  falkor_cruisename = "SCOPE_Falkor2"
  falkor_cruise_data <- get_cruise_by_name(falkor_cruisename)
  falkor_CMAP_SLA <- get_spacetime(tableName = 'tblAltimetry_REP',
                     varName = 'sla',
                     dt1=as.character(falkor_cruise_data$Start_Time),
                     dt2=as.character(falkor_cruise_data$End_Time),
                     lat1=falkor_cruise_data$Lat_Min,
                     lat2=falkor_cruise_data$Lat_Max,
                     lon1=falkor_cruise_data$Lon_Min,
                     lon2=falkor_cruise_data$Lon_Max)

  falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
  falkor_SLA <- falkor_ctddata %>%
    mutate(time=as.POSIXct(time)) %>%
    group_by(station) %>%
    summarize(lat=median(lat), lon=median(lon), time=median(time)) %>%
    difference_left_join(falkor_CMAP_SLA, max_dist = 0.125, by=c("lat", "lon")) %>%
    filter(as.Date(time.x)==as.Date(time.y)) %>%
    select(station, sla)

  write.csv(falkor_SLA, file="raw_data/falkor_SLA.csv", row.names = FALSE)
}
if(!file.exists("raw_data/meso_SLA.csv")){
  meso_SLA <- "http://scope.soest.hawaii.edu/data/mesoscope/km1709_CtdSla.txt" %>%
    read.table(skip = 1, col.names = c("station", "cast", "sla")) %>%
    mutate(sla=sla/100)
  write.csv(meso_SLA, file = "raw_data/meso_SLA.csv", row.names = FALSE)
}

# Combine and write out ----
falkor_SLA <- read.csv("raw_data/falkor_SLA.csv") %>%
  mutate(station=as.character(station)) %>%
  mutate(cast=1)
meso_SLA <- read.csv("raw_data/meso_SLA.csv")
dcm_depths <- read.csv("raw_data/all_DCMs.csv")
falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
meso_ctddata <- read.csv("raw_data/meso_SCOPEweb_ctd_summary.csv")

station_metadata <- meso_ctddata %>%
  mutate(cruise=ifelse(startsWith(station, "L")|station==15, "MC", "MT")) %>%
  rbind(falkor_ctddata %>% mutate(cruise="FK")) %>%
  left_join(dcm_depths, by=c("station", "cast")) %>%
  left_join(rbind(meso_SLA, falkor_SLA), by=c("station", "cast")) %>%
  filter(!(cruise=="FK" & station < 28)) %>%
  select(cruise, station, cast, time, lat, lon, DCM_depth, sla) %>%
  mutate(time=as.POSIXct(time, tz = "UTC")) %>% 
  mutate(local_hour=as.numeric(format(time, tz="HST", format = "%H")))

write.csv(station_metadata, file = "made_data/station_metadata.csv", row.names = FALSE)
```

```{r all_core and bottle_depth_data}
# Core Water Column Data from .gof file on the internet
if(!file.exists("raw_data/falkor_core.txt")){
  fk_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/fk180310.gof"
  download.file(fk_meta, destfile = "raw_data/falkor_core.txt")
}
if(!file.exists("raw_data/mesoscope_core.txt")){
  meso_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/mesoscope.gof"
  download.file(meso_meta, destfile = "raw_data/mesoscope_core.txt")
}
meso_core <- read_table("raw_data/mesoscope_core.txt", skip = 4, 
                        col_names = FALSE) %>%
  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")) %>%
  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  select(cruise, station, cast, rosette, pressure, 
         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
fk_core <- read_table("raw_data/falkor_core.txt", skip = 4, 
                        col_names = FALSE) %>%
  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")) %>%
  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
  mutate(cruise="FK") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  select(cruise, station, cast, rosette, pressure, 
         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
all_core <- bind_rows(fk_core, meso_core)
write.csv(all_core, "made_data/all_core.csv", row.names = FALSE)

bottle_depth_data <- all_core %>%
  distinct(cruise, station, cast, rosette, pressure)
write.csv(bottle_depth_data, "made_data/bottle_depth_data.csv", row.names = FALSE)
```

```{r plot station_metadata}
station_metadata <- read_csv("made_data/station_metadata.csv", show_col_types = FALSE)
stations_summarized <- station_metadata %>%
  group_by(cruise, station, cast) %>%
  summarize(lat=median(lat), lon=median(lon), time=median(time), sla=mean(sla)) %>%
  arrange(time)

station_labels <- stations_summarized %>%
  filter(station%in%c(seq(1, 15, 2), "L1", "L2", 62, 64, 77, 80)) %>%
  filter(cast==1)

gm <- get_map(location = c(-158, 24), zoom = 6)
map_gp <- ggmap(ggmap = gm) +
  geom_point(aes(x=lon, y=lat), size=4, color="#FFFFFFAA", data = station_metadata) +
  geom_path(aes(x=lon, y=lat), data = station_metadata) +
  geom_point(aes(x=lon, y=lat, color=sla), data = station_metadata) +
  geom_label_repel(aes(x=lon, y=lat, label=station), color="steelblue", 
                   data = station_labels, min.segment.length = 0) +
  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1) +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(-162, -155), ylim=c(21, 28))

sla_ts_gp <- station_metadata %>%
  ggplot() +
  geom_point(aes(x=time, y=sla, color=sla)) +
  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1) +
  theme_bw() +
  theme(legend.position = c(0.5, 0.55), axis.title.x = element_blank())
station_metadata_map <- grid.arrange(map_gp, sla_ts_gp)
ggsave(plot = station_metadata_map, filename = "station_metadata_map.png", 
       path = "figures", device = "png", width = 8, height = 6, dpi = 144, 
       type = "cairo")
```

## Other bottle data (cruise, station, cast, depth, PC, PN, PO4, N+N, DON, DOP, ATP, chl)

This data is mostly available on the SCOPE website via Excel files

```{r raw data download}
if(!file.exists("raw_data/Karl_FK_PCPN_WC_current.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
  falkor_CN_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/",
                "Data/Karl/Karl_FK_PCPN_WC_current.xlsx")
  download.file(falkor_CN_url, destfile = "raw_data/Karl_FK_PCPN_WC_current.xlsx", mode = "wb")
}
if(!file.exists("raw_data/Karl_FK_Nutrients_Final.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
  falkor_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "Karl/Karl_FK_Nutrients_Final.xlsx")
  download.file(falkor_nutrient_url, destfile = "raw_data/Karl_FK_Nutrients_Final.xlsx", mode = "wb")
}
if(!file.exists("raw_data/2017MS_CoreData_current.xls")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/2017MS_CoreData_current.xls
  transect_pcpn_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "SCOPEcore/2017MS_CoreData_current.xls")
  download.file(transect_pcpn_url, destfile = "raw_data/2017MS_CoreData_current.xls", mode = "wb")
}
if(!file.exists("raw_data/Karl_MS_Nutrient_Summary_KB.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_Nutrient_Summary_KB.xlsx
  transect_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/",
                "Karl_MS_Nutrient_Summary_KB.xlsx")
  download.file(transect_nutrient_url, destfile = "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx", mode = "wb")
}
if(!file.exists("raw_data/Karl_MS_pATP.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_pATP.xlsx
  atp_transect_url <- paste0("http://", scope_uname, ":", scope_pw,
         "@scope.soest.hawaii.edu/collaborators/datainventory/",
         "Data/Karl/Karl_MS_pATP.xlsx")
  download.file(atp_transect_url, destfile = "raw_data/Karl_MS_pATP.xlsx", mode = "wb")
}
```

```{r all_pcn}
fk_pcn <- "raw_data/Karl_FK_PCPN_WC_current.xlsx" %>%
  read_excel(skip=1) %>%
  slice(-1:-2) %>% 
  filter(is.na(Comments)) %>%
  mutate(station=as.numeric(str_extract(Name, "(?<=FK-)\\d+"))) %>%
  filter(station>=28) %>%
  mutate(rosette=as.numeric(str_extract(Name, "\\d+$"))) %>%
  mutate(cruise="FK", cast=1) %>%
  select(cruise, station, cast, rosette, abs_depth=`Depth (m)`, 
         PC_um=`C value (µmol/L)`, PN_um=`N value (µmol/L)`)
meso_pcn <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "PCPN") %>%
  slice(-1) %>%
  select(abs_depth=depth_sample, station, PC_um=particulate_carbon, 
         PN_um=particulate_nitrogen, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station %in% 16 ~ "L1",
    station == 17 ~ "L2"
  ))
all_pcn <- rbind(meso_pcn, fk_pcn)
write.csv(all_pcn, "made_data/all_pcn.csv", row.names = FALSE)
```

>>>>

library(tidyverse)
library(readxl)     # For "readxl"

@@

options(pillar.sigfig = 7)
```

## Download/collect raw data

Full list of files we need!
From SCOPE website:
  - FK_CTDsummary_current.txt
  - MS_CTDsummary_current.txt
  - Karl_FK_PCPN_WC_current.xlsx
  - Karl_FK_Nutrients_Final.xlsx
  - 2017MS_CoreData_current.xls
  - Karl_MS_Nutrient_Summary_KB.xlsx
  - Karl_MS_pATP.xlsx
  - Karl_FK_chlorophyll.xlsx
From email
  - FALKOR- Sample Log_INGALLS (From Tim)
  - MESO_SCOPE_Cruise_KM1709_Sample_Log (From Laura)
  - MESO-SCOPE_BeamAttenuation_forWill (From Benedetto, as are remainder)
  - MESO-SCOPE_DailyPAR_forWill
  - MESO-SCOPE_PercentPAR_forWill
  - kok1607_ctd_sla
  - km1709_ctd_sla
  - fk180310_ctd_sla
Manually created
  - hmdb_formulae.rds
  - MAA_masses.xlsx

```{r From SCOPE data inventory, eval = !file.exists("raw_data/FK_CTDsummary_current.txt")}
inventory_url <- paste0("http://", scope_uname, ":", scope_pw,
                        "@scope.soest.hawaii.edu/collaborators/datainventory/Data/")
from_scope_inventory <- c(
  "SCOPEcore/FK_CTDsummary_current.txt",
  "SCOPEcore/MS_CTDsummary_current.txt",
  "Karl/Karl_FK_PCPN_WC_current.xlsx",
  "Karl/Karl_FK_Nutrients_Final.xlsx",
  "SCOPEcore/2017MS_CoreData_current.xls",
  "Karl/Karl_MS_Nutrient_Summary_KB.xlsx",
  "Karl/Karl_MS_pATP.xlsx",
  "Karl/Karl_FK_Chlorophyll.xlsx"
)
sapply(from_scope_inventory, function(end_url){
  download.file(paste0(inventory_url, end_url), destfile = paste0("raw_data/", basename(end_url)), mode = "wb")
})
```

```{r From SCOPE FTP, eval=!file.exists("raw_data/mesoscope.gof")}
ftp_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/"
from_scope_ftp <- c("mesoscope.gof","fk180310.gof")
sapply(from_scope_ftp, function(end_url){
  download.file(paste0(ftp_url, end_url), destfile = paste0("raw_data/", basename(end_url)))
})
```

```{r From email}
# Go search for the files in email, we can't script this part
```

```{r Manually created}
# Go search for the files in Slack, we can't script this part
```

## SCOPE data cleaning and combining

```{r Sample log and DCM extraction}
col_types <- c("date", "text", "numeric", "text",
               "text", "date", "date", "text")
col_names <- c("date", "station", "cast", "sample_id", "depth",
               "start_time", "end_time", "volume")
falkor_log <- c("Anti-Cyclonic Eddy", "Cyclonic Eddy") %>%
  lapply(function(sheet_i){
    read_xlsx("raw_data/FALKOR- Sample Log_INGALLS.xlsx",
             sheet=sheet_i, skip = 5, 
             col_types = col_types, col_names = col_names) 
}) %>%
  bind_rows() %>%
  drop_na() %>%
  filter(sample_id!="1") %>%
  mutate(date=as.Date(date)) %>%
  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
  mutate(str_depth=ifelse(startsWith(depth, "DCM"), "DCM", paste0(depth, "m"))) %>%
  mutate(depth=ifelse(depth=="25", "Surface", "DCM")) %>%
  mutate(volume=as.numeric(str_extract(volume, "\\d+"))) %>%
  mutate(sample_id=paste0("S", station, "C", cast, "-", str_depth, "_", sample_id)) %>%
  select(-str_depth)

col_types <- c("text", "text", "text", "text",
               "text", "numeric", "date", "date",
               "text", "text", "skip", "skip", "skip")
col_names <- c("date", "station", "cast", "depth", "sample_id",
               "volume", "start_time", "end_time", "rosette", 
               "dissolved")
MT_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
          sheet = "Metabolites_and_Proteins", skip = 2,
          col_types = col_types, col_names=col_names) %>%
  drop_na() %>%
  mutate(date=ifelse(startsWith(date, "6/29"), "42916.0", date)) %>%
  mutate(date=as.Date(as.numeric(date), origin="1899-12-31")) %>%
  mutate(station=str_extract(station, "\\d+")) %>%
  mutate(cast=as.numeric(str_extract(cast, "\\d+"))) %>%
  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
  mutate(depth=case_when(
    depth=="15.0"~"Surface", depth=="175.0"~"Deep", TRUE~"DCM")
  ) %>%
  mutate(rosette=case_when(rosette=="?"~NA, rosette=="43026.0"~"10", rosette=="16- 24"~"16")) %>%
  mutate(rosette=as.numeric(rosette)) %>%
  # Fixing likely transcription error - sample looks like 15m samples and
  # all other triplicates were 19, 20, and 21
  mutate(rosette=ifelse(rosette==4 & cast==3 & station=="6", 21, rosette))

col_names <- c("station", "cast", "depth", "sample_id",
               "volume", "start_time", "end_time", "rosette", 
               "dissolved")
col_types <- c("text", "text", "text", "text",
               "numeric", "date", "date", "numeric", 
               "text", "skip")

MC_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
          sheet = "Metabolites_High_Resolution", skip = 2,
          col_types = col_types, col_names=col_names) %>%
  filter(!is.na(station)) %>%
  mutate(station=str_remove(station, "^S")) %>%
  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
  mutate(sample_id=str_replace(sample_id, "\\-", "less")) %>%
  mutate(sample_id=str_replace(sample_id, "\\+", "plus")) %>%
  mutate(depth_shift=str_extract(depth, "(\\-|\\+)\\d+")) %>%
  mutate(depth_shift=ifelse(is.na(depth_shift), 0, as.numeric(depth_shift))) %>%
  mutate(abs_depth=case_when(
    station=="L1"~108+depth_shift,
    station=="L2"~123+depth_shift
  )) %>%
  select(-depth_shift) %>%
  mutate(depth=str_replace(depth, "\\-", "less")) %>%
  mutate(depth=str_replace(depth, "\\+", "plus")) %>%
  mutate(rosette=as.numeric(rosette)) %>%
  mutate(sample_id=str_remove(sample_id, "^MHR")) %>%
  mutate(date=as.Date(case_when(
    station=="L1"~"2017-07-08",
    station=="L2"~"2017-07-09"
  )))

samp_log <- bind_rows(falkor_log, MC_log, MT_log)
dcm_depths <- samp_log %>%
  distinct(station, depth, abs_depth) %>%
  filter(depth=="DCM") %>%
  select(station, dcm_depth=abs_depth)
```

```{r station locations and CTD summaries}
col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
col_types <- cols(
  station = col_character(),
  cast = col_double(),
  month = col_character(),
  day = col_character(),
  year = col_double(),
  time = col_time(format = ""),
  lat_deg = col_double(),
  lat_min = col_double(),
  lat_dir = col_character(),
  lon_deg = col_double(),
  lon_min = col_double(),
  lon_dir = col_character(),
  max_depth = col_double(),
  bottles = col_double())
toDecDeg <- function(deg, min, dir){
  (deg+min/60)*ifelse(dir%in%c("S", "W"), -1, 1)
}
parseCTD <- function(file_path){
  file_path %>%
    read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}
meso_ctddata <- parseCTD("raw_data/MS_CTDsummary_current.txt")
falkor_ctddata <- parseCTD("raw_data/FK_CTDsummary_current.txt") %>%
  filter(time > "2018-03-23 04:01:51")
complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
```

```{r SLA parsing}
col_types <- cols(
  stn = col_character(),
  cast = col_double(),
  sla_corr = col_double()
)
falkor_sla <- read_csv("raw_data/fk180310_ctd_sla.txt", col_types = col_types) %>%
  mutate(cruise="FK")
meso_sla <- read_csv("raw_data/km1709_ctd_sla.txt", col_types = col_types) %>%
  mutate(cruise=ifelse(stn%in%c("15", "L1", "L2"), "MC", "MT"))
complete_sla <- bind_rows(falkor_sla, meso_sla) %>%
  rename(station="stn")
```

```{r Core water column parsing (.gof files)}
col_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(col_names, c(head(col_names, 6), tail(col_names, 4)))
gofReader <- function(filename){
  # From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
  # Quality Indicators:
  # Flag  Meaning
  # 1    not quality controled
  # 2    good data
  # 3    suspect (i.e.  questionable) data
  # 4    bad data
  # 5    missing data
  # 9    variable not measured during this cast
  raw_gof <- read_table(filename, skip = 4, col_names = col_names,
                        col_types = cols(.default = col_double()))
  qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
    unite("value", sep = "") %>%
    separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
    mutate(across(everything(), as.numeric))
  raw_gof[,qual_names][qual_df!=2] <- NA
  raw_gof %>% select(-matches("^q\\d+$"))
}
meso_core <- gofReader("raw_data/mesoscope.gof") %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  ))
falkor_core <- gofReader("raw_data/fk180310.gof") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  mutate(cruise="FK")
all_core <- bind_rows(meso_core, falkor_core)
```

```{r Parsing beam attenuation data}
beam_long <- read.csv("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv", 
                    skip = 3, header = FALSE) %>%
  `colnames<-`(paste(.[1,], .[2,], sep = " ")) %>%
  tail(-2) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(cols = -1, values_to = "beam_atten") %>%
  filter(!is.na(beam_atten)) %>%
  separate(name, into = c("station", "cast"), sep = " ") %>%
  rename(abs_depth=`NaN NaN`) %>%
  mutate(cast=as.numeric(cast))

# Convert to bottle locations
beam_core <- all_core %>%
  group_by(station, cast) %>%
  # filter(station==4, cast==2) %>%
  group_split() %>% #`[[`(1) -> core_subset
  map(function(core_subset){
    beam_subset <- beam_long %>%
      filter(station==unique(core_subset$station), cast==unique(core_subset$cast))
    if(nrow(beam_subset)==0)return(NULL)
    approxout <- with(beam_subset, approx(abs_depth, beam_atten, core_subset$pressure, ties = mean))
    core_subset %>%
      select(cruise, station, cast, rosette, pressure) %>%
      left_join(set_names(as.data.frame(approxout), c("pressure", "beam_atten")), 
                by="pressure", multiple="all")
  }, .progress = TRUE) %>%
  bind_rows() %>%
  distinct() %>%
  filter(beam_atten<0.2)
```

```{r PCPN parsing}
fk_pcn <- "raw_data/Karl_FK_PCPN_WC_current.xlsx" %>%
  read_excel(skip=1) %>%
  slice(-1:-2) %>% 
  filter(is.na(Comments)) %>%
  mutate(station=str_extract(Name, "(?<=FK-)\\d+")) %>%
  mutate(rosette=as.numeric(str_extract(Name, "\\d+$"))) %>%
  mutate(cruise="FK", cast=1) %>%
  select(cruise, station, cast, rosette, abs_depth=`Depth (m)`, 
         PC_um=`C value (µmol/L)`, PN_um=`N value (µmol/L)`)
meso_pcn <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "PCPN") %>%
  slice(-1) %>%
  select(abs_depth=depth_sample, station, PC_um=particulate_carbon, 
         PN_um=particulate_nitrogen, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station %in% 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  mutate(rosette=case_when(
    station=="L1"&cast==13&rosette==10~20, 
    station=="L1"&cast==6&rosette==15~16,
    station=="L1"&cast==6&rosette==14~15,
    TRUE~rosette))
all_pcn <- bind_rows(meso_pcn, fk_pcn)
```

```{r Nutrient parsing}
# Not really necessary because .gof file has nutrients in it already
# These files have more sigfigs but less depth accuracy (5, 25 instead of 4.7, 25.2)
# Actually, very necessary because the Falkor .gof file has N+N and PO4
# values switched??? Compare to Karl_FK_Nutrients_Final.xlsx
all_core %>%
  filter(cruise=="FK") %>%
  filter(!is.na(no23_um)) %>%
  filter(pressure>160) %>%
  select(station, cast, rosette, pressure, no23_um, po4_um)

falkor_nuts <- "raw_data/Karl_FK_Nutrients_Final.xlsx" %>%
  read_excel(skip = 3) %>%
  select(abs_depth=`Depth`, sample=`Sample ID`, po4_um=`µM PO4`, no23_um=`µM N+N`) %>%
  filter(!is.na(abs_depth)) %>%
  filter(!abs_depth=="Depth") %>%
  mutate(rosette=as.numeric(str_extract(sample, "\\d+$"))) %>% 
  mutate(sample=ifelse(startsWith(sample, "FK"), sample, NA)) %>%
  fill(sample) %>%
  mutate(station=str_extract(sample, "(?<=FK 180310[- ])\\d+")) %>%
  mutate(across(c(abs_depth, po4_um, no23_um, station), as.numeric)) %>%
  filter(station>=28) %>%
  mutate(station=as.character(station)) %>%
  select(-sample) %>%
  mutate(cast=1, cruise="FK") %>%
  arrange(station, abs_depth)

transect_profile_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Transect, L1,L2 profiles!A4:N180") %>%
  slice(-1:-2) %>%
  mutate(sd=as.numeric(sd)) %>%
  mutate(Si=case_when(
    Station=="8" & Cast==2 & Niskin==4 ~ NA,
    Station=="10" & Cast==1 & Niskin==3 ~ NA,
    Station=="14" & Cast==1 & Niskin==8 ~ NA,
    TRUE~Si
  )) %>%
  mutate(LLP=case_when(
    Station=="L2" & Cast==26 & Niskin==1 ~ NA,
    TRUE~LLP
  ))
highres_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 High res nuts!A4:P53") %>%
  slice(-1:-2) %>%
  filter(!is.na(Station))
diel_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Diel nuts!A4:O76") %>%
  slice(-1:-3) %>%
  filter(!is.na(Station)) %>%
  mutate(LLP=case_when(
    Station=="L1"&Cast==11&Niskin...3==20~NA,
    Station=="L1"&Cast==19&Niskin...3==9~NA,
    TRUE~LLP
  )) %>%
  mutate(Si=case_when(
    Station=="L1" & Cast==10 & Niskin...3==10 ~ NA,
    Station=="L1" & Cast==15 & Niskin...3==10 ~ NA,
    Station=="L1" & Cast==17 & Niskin...3==11 ~ NA,
    Station=="L2" & Cast==6 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==12 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==17 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==20 & Niskin...3==18 ~ NA,
    Station=="L2" & Cast==25 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==11 & Niskin...3==3 ~ NA,
    TRUE~Si
  ))
